Analysis of the 2005--\myrevision{2016} Earthquake Sequence in Northern Iran Using the Visibility Graph Method

### 2017-06-23 abstract.tex
Suggested changes by Luciano added.

%
We present an analysis of the seismicity of northern Iran in the period between 2005 and \myrevision{2016} using a recently introduced method based on concepts of graph theory. The method relies on the inter-event visibility defined in terms of a connectivity degree parameter, $k$, which is correlated with the earthquake magnitude, $M$. Previous studies show that the slope of the $k$-$M$ correlation, $m$, also observes a relationship with the $b$-value from the Gutenberg-Richter law, thus rendering the graph analysis useful to examine the seismicity of a region. \myrevision{These correlations seem to hold for the analysis of relatively small sequences of earthquakes, offering the possibility of studying seismicity parameters in time}. We apply this approach to the case of the seismicity of northern Iran, using an earthquake catalog for the tectonic seismic regions of Azerbaijan, Alborz, and Kopeh Dagh. We use results drawn for this region with the visibility graph approach in combination with results from other similar studies to further improve the universal relationship between $m$ and $b$, and show that the visibility graph approach can be considered as a valid alternative for analyzing regional seismicity properties and earthquake sequences.

>>>>

%
We present an analysis of the seismicity of northern Iran in the period between 2005 and \myrevision{2016} using a recently introduced method based on concepts of graph theory. The method relies on the inter-event visibility defined in terms of a connectivity degree parameter, $k$, which is correlated with the earthquake magnitude, $M$. \myrevision{Previous studies show that the slope $m$ of the line fitting the  $k-M$  plot by the least square method}, also observes a relationship with the $b$-value from the Gutenberg-Richter law, thus rendering the graph analysis useful to examine the seismicity of a region. \myrevision{These correlations seem to hold for the analysis of relatively small sequences of earthquakes, offering the possibility of studying seismicity parameters in time}. We apply this approach to the case of the seismicity of northern Iran, using an earthquake catalog for the tectonic seismic regions of Azerbaijan, Alborz, and Kopeh Dagh. We use results drawn for this region with the visibility graph approach in combination with results from other similar studies to further improve the universal relationship between $m$ and $b$, and show that the visibility graph approach can be considered as a valid alternative for analyzing regional seismicity properties and earthquake sequences.

### 2017-06-23 seismicity.tex
full table 1
Suggested changes by Luciano added.

\myrevision{The whole dataset we downloaded from IIEES can be seen in Fig.~\ref{fig:seismicity}a, which shows events scaled by their magnitude $M_w$ at their epicenter location for the three seismic zones of interest. We refer to the catalog composed by the events shown in this figure as the \textit{whole} catalog. We then} determine the minimum magnitude of completeness, $M_c$, for each region\myrevision{, and based on the value of $M_c$, we filter out the smaller events and obtain a reduced dataset of events to which we refer to as the \textit{complete} catalog.}

@@

\myrevision{Subsequently, we} obtained the $b$-value in the Gutenberg-Richter law for each region following the maximum likelihood estimation \citep{Aki1965}, in which
% 
\begin{equation}
\myrevision{
	b = \frac
		{ log_{10}(e) }
		{ \overline{M} - \left( M_{\min} - \frac{\Delta M_{\mathrm{bin}}}{2} \right) } \, ,
}
\end{equation}
%
\noindent%
where $e$ is the mathematical constant or Euler's number, $\overline{M}$ is the average magnitude, \myrevision{$M_{\min}$ is the minimum magnitude in the sample, and $\Delta M_{\mathrm{bin}}$ is the bin size used to discretize the magnitude scale}. Here, the value of $M_{\min}$ is taken as the minimum completeness magnitude $M_c$ obtained or selected for each region's sub-catalog. \myrevision{In this study we assumed $\Delta M_{\mathrm{bin}} = 0.1$}.

@@

\myrevision{
\begin{table*}
	\caption{Total number of earthquakes \myrevision{for the whole, complete, and declustered catalogs}, and seismic parameters $M_c$ and $b$ \myrevision{(including standard deviation) for the complete catalogs} of the three seismic zones of northern Iran.}
	\centering\small
	\myrevision{
	\begin{tabular}{lcccccc}
		\hline
		Region       & \multicolumn{3}{c}{Number of Events in Catalog} & $M_c$ & \multicolumn{2}{c}{$b$-value} \\
		\cline{2-4}\cline{6-7} 
		             & Whole & Complete & Declustered &     & Complete             & Declustered \\
		\hline
		Azerbaijan   & 1047  &  202     &     ?       & 3.5 & 0.85491 $\pm$ 0.0506 & ? \\
		Alborz       & 1898  &  575     &     ?       & 3.0 & 0.81578 $\pm$ 0.0279 & ? \\
		Kopeh Dagh   &  640  &  117     &     ?       & 3.5 & 0.78297 $\pm$ 0.0494 & ? \\
		\hline
	\end{tabular}
	}
	\label{tab:seismicity}
\end{table*}
}

\begin{figure*}%[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-05} 
	\caption{Time-magnitude sequences for the northern Iran seismic zones of Azerbaijan, Alborz and Kopeh Dagh, during the period between 1 January 2005 and 31 December \myrevision{2016}. The event occurrences are represented by vertical bars or sticks of length equal to the $M_w$ magnitude of each earthquake. These sequences correspond to the \myrevision{whole} catalog but in each case we consider only events with $M_w \geq M_c$. As shown in Fig.~\ref{fig:vg}, the dots at the top of each stick represent the nodes of the visibility graphs. The edges or links are omitted for visual convenience. Highlighted in the figure are the largest events in each sub-catalog, namely the 2012 $M_w$ 6.4 East Azerbaijan, 2010 $M_w$ 5.8 Damghan, and 2012 $M_w$ 5.4 Neyshabur earthquakes.}
	\label{fig:mag-time}
\end{figure*}

>>>>

\myrevision{The whole dataset we downloaded from IIEES can be seen in Fig.~\ref{fig:seismicity}a, which shows events scaled by their magnitude $M_w$ at their epicenter location for the three seismic zones of interest. We refer to the catalog composed by the events shown in this figure as the \textit{whole} catalog. We then} determine the magnitude of completeness, $M_c$, for each region\myrevision{, and based on the value of $M_c$, we filter out the smaller events and obtain a reduced dataset of events to which we refer to as the \textit{complete} catalog.}

@@

\myrevision{Subsequently, we} obtained the $b$-value in the Gutenberg-Richter law for each region following the maximum likelihood estimation \citep{Aki1965}, in which
% 
\begin{equation}
\myrevision{
	b = \frac
		{ log_{10}(e) }
		{ \overline{M} - \left( M_{\min} - \frac{\Delta M_{\mathrm{bin}}}{2} \right) } \, ,
}
\end{equation}
%
\noindent%
where $e$ is the mathematical constant or Euler's number, $\overline{M}$ is the average magnitude, \myrevision{$M_{\min}$ is the minimum magnitude in the sample, and $\Delta M_{\mathrm{bin}}$ is the bin size used to discretize the magnitude scale}. Here, the value of $M_{\min}$ is taken as the completeness magnitude $M_c$ obtained or selected for each region's sub-catalog. \myrevision{In this study we assumed $\Delta M_{\mathrm{bin}} = 0.1$}.

@@

\myrevision{
\begin{table*}
	\caption{Total number of earthquakes \myrevision{for the whole, complete, and declustered catalogs}, and seismic parameters $M_c$ and $b$ \myrevision{(including standard deviation) for the complete catalogs} of the three seismic zones of northern Iran.}
	\centering\small
	\myrevision{
	\begin{tabular}{lcccccc}
		\hline
		Region       & \multicolumn{3}{c}{Number of Events in Catalog} & $M_c$ & \multicolumn{2}{c}{$b$-value} \\
		\cline{2-4}\cline{6-7} 
		             & Whole & Complete & Declustered &     & Complete        & Declustered \\
		\hline
		Azerbaijan   & 1047  &  202     &   102       & 3.5 & 0.85 $\pm$ 0.05 & 0.88 $\pm$ 0.07 \\
		Alborz       & 1898  &  575     &   440       & 3.0 & 0.82 $\pm$ 0.03 & 0.79 $\pm$ 0.03 \\
		Kopeh Dagh   &  640  &  117     &    96       & 3.5 & 0.78 $\pm$ 0.05 & 0.73 $\pm$ 0.05 \\
		\hline
	\end{tabular}
	}
	\label{tab:seismicity}
\end{table*}
}

\begin{figure*}%[t]
	\centering
	\includegraphics[width=\textwidth]{figures/pdf/figure-05} 
	\caption{Time-magnitude sequences for the northern Iran seismic zones of Azerbaijan, Alborz and Kopeh Dagh, during the period between 1 January 2005 and 31 December \myrevision{2016}. The event occurrences are represented by vertical bars or sticks of length equal to the $M_w$ magnitude of each earthquake. These sequences correspond to the \myrevision{complete} catalog. As shown in Fig.~\ref{fig:vg}, the dots at the top of each stick represent the nodes of the visibility graphs. The edges or links are omitted for visual convenience. Highlighted in the figure are the largest events in each sub-catalog, namely the 2012 $M_w$ 6.4 East Azerbaijan, 2010 $M_w$ 5.8 Damghan, and 2012 $M_w$ 5.4 Neyshabur earthquakes.}
	\label{fig:mag-time}
\end{figure*}
